Retrofitting Effect Handlers onto OCaml

### 2020-11-17 retro-concurrency.tex
Updates

@@

      \textsc{Handle}   & (\handle{e}{h}, \env, k, \cstack) & \ostep
                        & \configa{e}{\env}{\ostacka{([],(h,\env)) \kcons k}{\cstack}} \\
      \textsc{ExHn}     & (v, \_, (\ffun{\exnval{l}}::\_,(h,\env)) \kcons k', \cstack) & \ostep
                        & \configa{e}{\envext{\env}{x}{v}}{\ostacka{k'}{\cstack}}
                        & \text{if } \{\caseexn{l}{x}{e}\} \in h \\
      \textsc{ExFwdC}   & (v, \env, [\ffun{\exnval{l}}::\_,(h,\_)], \cstacka{\fl'}{\ostack}) & \ostep
                        & \configa{v}{\env}{\cstacka{\ffun{\exnval{l}}::\fl'}{\ostack}}
                        & \text{if } \{\caseexn{l}{\_}{\_}\} \notin h \\
      \textsc{ExFwdFib} & (v, \env, (\ffun{\exnval{l}}::\_,(h,\_)) \kcons (\fl',\hc') \kcons k', \cstack) & \ostep
                        & \configa{v}{\env}{\ostacka{(\ffun{\exnval{l}}::\fl',\hc') \kcons k'}{\cstack}}
                        & \text{if } \{\caseexn{l}{\_}{\_}\} \notin h \\

@@

\textbf{\textit{Top-level reductions.}} The initial configuration for an expression
$e$ is $\configa{e}{\emptyset}{\cstacka{[]}{\bullet}}$, where the environment
and the stack are empty. The top-level reductions (Figure ~\ref{sem:toplevel})
can be performed by either by taking a C step $\cstep$ or an OCaml step
$\ostep$.

\textbf{\textit{C reductions.}} We can take a C step (Figure~\ref{sem:cstep}) by
taking an administrative reduction step $\rightsquigarrow$. The administrative
reductions are common to both C and OCaml. The rules \textsc{Var},
\textsc{Arith1}, \textsc{Arith2}, \textsc{App1}, \textsc{App2} and
\textsc{App3} are standard. \textsc{Arith3} performs the arithmetic operation
on the integers ($\llbracket n_1 \odot n_2 \rrbracket$). \textsc{Raise} pushes
an function frame with exception value to indicate that an exception is being
raised. Similarly, \textsc{Perform} pushes a function frame with an effect
value with an empty continuation $[[],(\{\caseval{x}{x}\},\emptyset)]$ with no
captured frames and an empty handler with an identity return case alone. We
shall return to \textsc{Resume1} and \textsc{Resume2} later.

@@

\textbf{\textit{OCaml reductions.}} In OCaml (Figure~\ref{sem:ostep}),
reductions always occur on the top-most fiber in the current stack.
\textsc{AdminO} performs administrative reductions. \textsc{CallO} evaluates an
OCaml function on the current stack. \textsc{ExtCall} captures the behaviour of
external calls, which are evaluated on an empty C stack with the current OCaml
stack as its tail. \textsc{RetToC} returns a value to the enclosing C stack. In
this case, we have exactly one fiber on the stack, and this was created in the
rule \textsc{Callback}, whose handler has identity return case alone and the
environment is empty. \textsc{RetFib} returns value from a fiber to the
previous one, evaluating the body of the return case.

The rule \textsc{Handle} installs a handler by pushing a fiber with no frames
and the given handler. The rule \textsc{ExHn} handles an exception, if the
current handler has a matching exception case, unwinding the current fiber. The
rule \textsc{ExFwdC} forwards the exception to C. Here, there is exactly one
fiber on the current stack, and the handler does not have a matching exception
case, which we know is the case (see \textsc{Callback} rule). The rule
\textsc{ExFwdFib} forwards the exception to the next fiber if the current
handler does not handle it.

>>>>

\vspace{-2mm}

@@

									& & & & \text{and } k \neq [] \\

      \textsc{Handle}   & (\handle{e}{h}, \env, k, \cstack) & \ostep
                        & \configa{e}{\env}{\ostacka{([],(h,\env)) \kcons k}{\cstack}} \\
      \textsc{ExnHn}     & (v, \_, (\ffun{\exnval{l}}::\_,(h,\env)) \kcons k', \cstack) & \ostep
                        & \configa{e}{\envext{\env}{x}{v}}{\ostacka{k'}{\cstack}}
                        & \text{if } \{\caseexn{l}{x}{e}\} \in h \\
      \textsc{ExnFwdC}   & (v, \env, [\ffun{\exnval{l}}::\_,(h,\_)], \cstacka{\fl'}{\ostack}) & \ostep
                        & \configa{v}{\env}{\cstacka{\ffun{\exnval{l}}::\fl'}{\ostack}}
                        & \text{if } \{\caseexn{l}{\_}{\_}\} \notin h \\
      \textsc{ExnFwdFib} & (v, \env, (\ffun{\exnval{l}}::\_,(h,\_)) \kcons (\fl',\hc') \kcons k', \cstack) & \ostep
                        & \configa{v}{\env}{\ostacka{(\ffun{\exnval{l}}::\fl',\hc') \kcons k'}{\cstack}}
                        & \text{if } \{\caseexn{l}{\_}{\_}\} \notin h \\

@@

\subsubsection{Top-level reductions}

The initial configuration for an expression $e$ is
$\configa{e}{\emptyset}{\cstacka{[]}{\bullet}}$, where the environment and the
stack are empty. The top-level reductions (Figure ~\ref{sem:toplevel}) can be
performed by either by taking a C step $\cstep$ or an OCaml step $\ostep$.

\subsubsection{C reductions}

We can take a C step (Figure~\ref{sem:cstep}) by taking an administrative
reduction step $\rightsquigarrow$. The administrative reductions are common to
both C and OCaml. The rules \textsc{Var}, \textsc{Arith1}, \textsc{Arith2},
\textsc{App1}, \textsc{App2} and \textsc{App3} are standard. \textsc{Arith3}
performs the arithmetic operation on the integers ($\llbracket n_1 \odot n_2
\rrbracket$). \textsc{Raise} pushes an function frame with exception value to
indicate that an exception is being raised. Similarly, \textsc{Perform} pushes
a function frame with an effect value with an empty continuation
$[[],(\{\caseval{x}{x}\},\emptyset)]$ with no captured frames and an empty
handler with an identity return case alone. We shall return to \textsc{Resume1}
and \textsc{Resume2} later.

@@

\subsubsection{OCaml reductions}

In OCaml (Figure~\ref{sem:ostep}), reductions always occur on the top-most
fiber in the current stack. \textsc{AdminO} performs administrative reductions.
\textsc{CallO} evaluates an OCaml function on the current stack.
\textsc{ExtCall} captures the behaviour of external calls, which are evaluated
on an empty C stack with the current OCaml stack as its tail. \textsc{RetToC}
returns a value to the enclosing C stack. In this case, we have exactly one
fiber on the stack, and this was created in the rule \textsc{Callback}, whose
handler has identity return case alone and the environment is empty.
\textsc{RetFib} returns value from a fiber to the previous one, evaluating the
body of the return case.

The rule \textsc{Handle} installs a handler by pushing a fiber with no frames
and the given handler. The rule \textsc{ExnHn} handles an exception, if the
current handler has a matching exception case, unwinding the current fiber. The
rule \textsc{ExnFwdC} forwards the exception to C. Here, there is exactly one
fiber on the current stack, and the handler does not have a matching exception
case, which we know is the case (see \textsc{Callback} rule). The rule
\textsc{ExnFwdFib} forwards the exception to the next fiber if the current
handler does not handle it.
